Crunching Large Graphs with Commodity Processors

### 2011-01-26 doc/hotpar2011/softxmt-hotpar2011.tex
Reword introduction

\begin{abstract}
   Crunching large graphs is the basis of many emerging applications,
   such as social network analysis and bioinformatics. Graph analytics
   algorithms exhibit little locality and therefore present significant
   performance challenges. Hardware multithreading systems (e.g, Cray
   XMT) have proven that with enough concurrency, we can tolerate long
   latencies. Unfortunately, this solution is not available with
   commodity parts.

@@

Many important emerging applications such as social network analysis,
bioinformatics and analyzing vast sensor data from actual networks
(e.g, water, electrical grid) are based on crunching very large
graphs. Unfortunately, the computational cost of these applications
gets quickly out of hand. For example, in January 2011, Facebook had
500 million active users with an average of 130 friends each
\cite{Facebook:2011p91}. Its friend suggestion algorithm in 2010 ran on
  a 40-node cluster with 72GB memory per node, and, still, new
  suggestions could only be computed every two days
  \cite{Backstrom:2010p90}. Speeding up these applications at a low
  cost would have a large impact in how we deal and react to the fast
  growth of the data-set and would make the results even more
  valuable.

@@

We believe we can build a system based mostly on commodity parts that
can attain XMT-like performance with a familiar, also XMT-like,
programming model, but at a fraction of the cost. Therefore, our goal
is to build a system that has good performance on low-locality graph
codes but is implemented with cheap commodity processors and simple
additional support in an FPGA. This approach also has the added
benefit of not sacrificing general-purpose performance.

@@

%We minimize context size by treating context switches as function
%calls, as in~\cite{charm}. This allows us 
%to save and restore the
%minimum number of registers allowed by the ABI, and 
%Simon:  not really true -- we could save fewer if we knew which of the
%callee save registers weren't going to be clobbered.
We treat context switches as function calls, as in~\cite{charm}. This allows us 
to depend on the compiler to save and restore caller-save registers while we explicitly save and restore all the callee-save registers:  the size of a coroutine context is hardly more than the stack space needed to save these registers.  We minimize
context switch time by doing all switching and scheduling in user space (kernel traps are at least an order of magnitude slower.)

>>>>

\begin{abstract}
   Crunching large graphs is the basis of many emerging applications,
   such as social network analysis and bioinformatics. Graph analytics
   algorithms exhibit little locality and therefore present significant
   performance challenges. Hardware multithreading systems (e.g, Cray
   XMT) show that with enough concurrency, we can tolerate long
   latencies. Unfortunately, this solution is not available with
   commodity parts.

@@

Many important emerging applications such as social network analysis,
bioinformatics, and sensor networks rely on crunching very large
graphs. Unfortunately, the computational cost of these applications
gets quickly out of hand. For example, in January 2011, Facebook had
500 million active users with an average of 130 friends each
\cite{Facebook:2011p91}. Its friend suggestion algorithm in 2010 ran on
  a 40-node cluster with 72GB memory per node, and, still, new
  suggestions could only be computed every two days
  \cite{Backstrom:2010p90}. Speeding up these applications at a low
  cost would have a large impact in how we deal with large data-sets and make them even more valuable.

@@

We believe we can build a system based mostly on commodity parts that
can attain XMT-like performance with a familiar, XMT-like,
programming model, but at a fraction of the cost. Therefore, our goal
is to build a system that has good performance on low-locality graph
codes but is implemented with cheap commodity processors and simple
additional support in an FPGA. This approach also has the added
benefit of not sacrificing general-purpose performance.

@@

We minimize context size by treating context switches as function
calls, as in~\cite{charm}. This allows us to save and restore the
minimum number of registers allowed by the ABI, and depend on the
compiler to save and restore any other registers in use. We minimize
context switch time by doing all switching and scheduling in user space (kernel traps are at least an order of magnitude slower.)
